---
title: "SIM Project 1"
author: "Adria Casanova", "Victor Garcia", "Zhengyong Ji"
date: "November, 19th 2023"
output: 
  pdf_document: 
    toc: true
    toc_depth: 3
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# The file contains some numeric variables (retain all of them) and many factors (restrict to 10 to reduce the effort).


```{r,echo=FALSE}
# Load the data
if(!is.null(dev.list())) dev.off()
rm(list = ls())

df_test = load()
df_train = load()
```


```{r}
# Import the necesary library

library(car)
library(missMDA)
```


**1. Exploratory Data Analysis and Model Fitting should take train sample only.**

```{r}



```


**2. Create factors for retained qualitative variables. Train and Test samples.**

```{r}

```


**3. Determine if the response variable (charges) has an acceptably normal distribution.**

```{r}

```


**4. Address tests to discard serial correlation.**

```{r}

```


**5. Detect univariant and multivariant outliers, errors and missing values (if any) and apply an imputation technique if needed.**

```{r}

```


**6. Preliminary exploratory analysis to describe observed relations has to be undertaken.**

```{r}

```


**7. If you can improve linear relations or limit the effect of influential data, you must consider suitable transformations for variables.**

```{r}

```


**8. Apart from the retained factor variables, you can consider other categorical variables that can be defined from categorized numeric variables. Do not forget to implement new variable definitions in the test sample.**

```{r}

```


**9. You must take into account possible interactions between categorical and numerical variables.**

```{r}

```


**10. When building the model, you should study the presence of multi-collinearity and try to reduce their impact on the model for easier interpretation. **

```{r}

```


**11. You should build the model using a technique for selecting variables (removing no significant predictors and/or stepwise selection of the best models).**

```{r}

```


**12.The validation of the model has to be done with graphs and / or suitable tests to verify model assumptions.**

```{r}

```


**13. You must include the study of unusual and / or influential data. **

```{r}

```


**14. The resulting model should be interpreted in terms of the relationships of selected predictors and its effect on the response variable. **

```{r}

```


**15. You have to apply your final model to the test sample and roughly assess forecasting capability.**

```{r}

```

